Quantifying Spatially Resolved Hydration Thermodynamics Using Grid Inhomogeneous Solvation Theory [Article v\versionnumber]

### 2022-12-19 manuscript/manuscript.tex

An example Amber input file for GIST might look like this:
\begin{lstlisting}[style=amber-in]
	restrained 100 ns NpT
	&cntrl
	ntx=5, irest=1,
	ioutfm=1,
	ntb=2, iwrap=1,
	ntr=1, restraint_wt=100.0,
	restraintmask="!@H=&!:WAT",
	ntp=1, pres0=1.0, taup=1.0,
	ntc=2, ntf=2,
	ntt=3, tempi=300.0, temp0=300.0, gamma_ln=2,
	nstlim=50000000, dt=0.002,
	ntwr=50000, ntwx=5000, ntpr=5000,
	/
\end{lstlisting}
This file will run a 100 ns NpT MD simulation while keeping all atoms except for hydrogens and waters restrained. Note that these are very standard MD settings except for the restraints.

Run the MD using \software{pmemd.cuda}:
\begin{lstlisting}[style=bash]
	pmemd.cuda \
	-O \
	-i 100ns-npt-restraint.in \
	-o md-01.out \
	-p solvated.parm7 \
	-c EQUIL-DONE.rst \
	-x md-01.nc \
	-r md-01.ncrst \
	-ref EQUIL-DONE.rst
\end{lstlisting}
\subsubsection{GIST analysis}
To start our GIST analysis of biotin, we'll first have to decide on the parameters defining the GIST grid during the analyis.
We recommend that you center the biotin molecule in space:
\begin{lstlisting}[style=cpptraj]
	parm solvated.parm7
	trajin md-01.nc 1 1 1
	autoimage !(:WAT) origin
	strip :WAT
	trajout biotin-centered.pdb
\end{lstlisting}

To decide on a grid size, find the extent (minimum and maximum) of the x-, \mbox{y-,} and z-coordinates of the molecule.
This can be done e.g., using the \software{mdtraj} package in Python.
\begin{lstlisting}[style=python]
	import mdtraj as md
	pdb = md.load("biotin-centered.pdb")
	# mdtraj converts to nm.
	# *10 to convert back to Angstrom.
	min_xyz = pdb.xyz[0].min(0) * 10.
	max_xyz = pdb.xyz[0].max(0) * 10.
	print(*list(zip("xyz", min_xyz, max_xyz)), sep="\n")
\end{lstlisting}
For a system centered at the origin with a spacing of \SI{0.5}{\angstrom} and a buffer of, e.g., \SI{15}{\angstrom} between the solute and the grid wall, you need $4.0\times (d+15)$ bins in the respective direction, where $d$ is the larger of $-$min and $+$max in the respective direction.
To get bulk-like water properties at the grid border, a buffer distance of \SI{15.0}{\angstrom} is reasonable for biotin.

@@

\begin{lstlisting}[style=cpptraj]
	parm solvated.parm7
	trajin md-01.nc
	autoimage !(:WAT) origin
	# optionally remove the "pme" option
	gist griddim 100 100 100 out gist.dat \
	refdens 0.03287 pme
\end{lstlisting}
\subsubsection{Interpreting and visualizing the results}
\todo{write this, show the various GIST properties here: Eww, Esw, S.}
\subsection{Applying GIST for the Streptavidin-Biotin complex}
In this section we will apply GIST to the full complex and discuss each step of the process in more detail to outline pitfalls. We will afterwards introduce advanced analysis workflows which allow you to modify and better visualize the resulting data.
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

>>>>

An example Amber input file for GIST might look like this:
\begin{lstlisting}[style=amber-in]
	restrained 100 ns NpT
	&cntrl
	ntx=5, irest=1,
	ioutfm=1,
	ntb=2, iwrap=1,
	ntr=1, restraint_wt=100.0,
	restraintmask="!@H=&!:WAT",
	ntp=1, pres0=1.0, taup=1.0,
	ntc=2, ntf=2,
	ntt=3, tempi=300.0, temp0=300.0, gamma_ln=2,
	nstlim=50000000, dt=0.002,
	ntwr=50000, ntwx=5000, ntpr=5000,
	/
\end{lstlisting}

This file will run a 100 ns NpT MD simulation while keeping all atoms except for hydrogens and waters restrained. Note that these are very standard MD settings except for the restraints.

Run the MD using \software{pmemd.cuda}:
\begin{lstlisting}[style=bash]
	pmemd.cuda \
	-O \
	-i 100ns-npt-restraint.in \
	-o md-01.out \
	-p solvated.parm7 \
	-c EQUIL-DONE.rst \
	-x md-01.nc \
	-r md-01.ncrst \
	-ref EQUIL-DONE.rst
\end{lstlisting}
\subsubsection{GIST analysis}
To start our GIST analysis of biotin, we'll first have to decide on the parameters defining the GIST grid during the analyis.
We recommend that you center the biotin molecule in space:
\begin{lstlisting}[style=cpptraj]
	parm solvated.parm7
	trajin md-01.nc 1 1 1
	autoimage !(:WAT) origin
	strip :WAT
	trajout biotin-centered.pdb
\end{lstlisting}
To decide on a grid size, find the extent (minimum and maximum) of the x-, \mbox{y-,} and z-coordinates of the molecule.
This can be done e.g., using the \software{mdtraj} package in Python.
\begin{lstlisting}[style=python]
	import mdtraj as md
	pdb = md.load("biotin-centered.pdb")
	# mdtraj converts to nm.
	# *10 to convert back to Angstrom.
	min_xyz = pdb.xyz[0].min(0) * 10.
	max_xyz = pdb.xyz[0].max(0) * 10.
	print(*list(zip("xyz", min_xyz, max_xyz)), sep="\n")
\end{lstlisting}
For a system centered at the origin with a spacing of \SI{0.5}{\angstrom} and a buffer of, e.g., \SI{15}{\angstrom} between the solute and the grid wall, you need $4.0\times (d+15)$ bins in the respective direction, where $d$ is the larger of $-$min and $+$max in the respective direction.
To get bulk-like water properties at the grid border, a buffer distance of \SI{15.0}{\angstrom} is reasonable for biotin.

@@

\begin{lstlisting}[style=cpptraj]
	parm solvated.parm7
	trajin md-01.nc
	autoimage !(:WAT) origin
	# optionally remove the "pme" option
	gist griddim 100 100 100 out gist.dat \
	refdens 0.03287 pme
\end{lstlisting}
\subsubsection{Interpreting and visualizing the results}

\todo{write this, show the various GIST properties here: Eww, Esw, S. Show how to calculate A and E with gisttools}

\subsection{Applying GIST for the Streptavidin-Biotin complex}
In this section we will apply GIST to the full complex and discuss each step of the process in more detail to outline pitfalls. We will afterwards introduce advanced analysis workflows which allow you to modify and better visualize the resulting data.
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.
